---
title: "Exemplary Code Chunk"
author: "Maveen Mallah"
date: "2024-05-01"
output: pdf_document
---

```{r, include=FALSE}
knitr::opts_chunk$set(echo = FALSE,
                      warning = FALSE,
                      tidy = FALSE,
                      message = FALSE,
                      fig.align = 'center',
                      out.width = "100%")
options(knitr.table.format = "html") 
```

```{r, include=FALSE}
knitr::write_bib(c('posterdown', 'rmarkdown','pagedown'), 'packages.bib')
```

```{r}
library(readr)
library(tidyverse)
library(posterdown)
library(ggplot2)
library(dplyr)
library(tidyr)
library(knitr)
library(kableExtra)
library(stargazer)
library(modelsummary)
library(gt)
```

```{r, include=FALSE}
data = read_csv('https://www.qogdata.pol.gu.se/data/qog_oecd_cs_jan24.csv')
sldata = data |> 
  select(cname, gggi_eas, gggi_ggi, gggi_hss, gggi_pes, gggi_pos, gii_gii) |> 
  na.omit()
```

```{r, include =TRUE}
graph1 = ggplot(sldata, aes(x = gii_gii, y = gggi_ggi)) +
  geom_point(color = "powderblue", size = 3) +  
  geom_smooth(method = "lm", se = FALSE, color = "navy") + #linear regression line 
  labs(title = "Comparison of Gender Inequality Index and Global Gender Gap Index",
       x = "Gender Inequality Index",
       y = "Global Gender Gap Index") +
theme_minimal() +
theme(axis.text.y = element_text(size = 8, color = "black", family = "Times"), #Times New Roman Font 
      axis.text.x = element_text(size = 8, color = "black", family = "Times"),
      plot.title = element_text(hjust = 0.5, color = "black", family = "Times"), 
      axis.title.y = element_text(color = "black", family = "Times"),  
      axis.title.x = element_text(color = "black", family = "Times"),  
      axis.line.x = element_line(color = "black"),#Add black line 
      axis.line.y = element_line(color = "black"), #Add black line 
      axis.ticks.x = element_line(color = "black"), 
      axis.ticks.length.x = unit(0.1, "cm"),  
      axis.ticks.y = element_line(color = "black"), 
      axis.ticks.length.y = unit(0.1, "cm"),
      panel.grid.major = element_blank(),  
      panel.grid.minor = element_blank()) 

print(graph1)
