---
title: "Unveiling Societal Strife: Analyzing Urban Social Disorder"
subtitle: "SIS750:Data Analysis"
author: "Maveen Mallah, M.A."
institute: "mm3664a@american.edu"
output:
  beamer_presentation:
    theme: "Hannover"
    colortheme: "lily"
fontsize: 10pt
classoption: "aspectratio=169"

---

```{r setup, include=FALSE}
## Libraries
  library(knitr)
  library(rmdformats)
  library(tidyverse)
  library(haven)
  library(readxl)
  library(patchwork)
  library(cowplot)
  library(dplyr)
  library(knitr)
  library(ggplot2)


## Global options
knitr::opts_chunk$set(
  echo = FALSE, 
  prompt = FALSE,
  tidy = FALSE, 
  comment = NA,
  message = FALSE, 
  warning = FALSE,
  fig.align = 'center')
```

## Data Selection
- Urban Social Disorder Version 3.0 (2022) provides a comprehensive overview of societal conflict dynamics across various regions.

- The dataset comprises events occurring within various cities and underwent processing to aggregate event types and participation counts across multiple cities within each country. Event types and participation counts from multiple cities within each country were averaged together.

- **PTYPE** categorizes societal conflicts based on their overall scale and nature. The graph helps visualize the distribution of conflicts across various categories. 

- **NPART **categorizes the participants involved in societal conflicts. It tracks two metrics ***participants**: everyone actively involved in the event regardless of side* and ***affected people:** those directly impacted by the event, even if they weren't actively participating*. 


## Analysis 1: PTYPE
```{r, out.width= "70%", out.height= "83%"} 
#size adjustment 
events <- read_excel("events.xlsx")
frequency_table1 = table(events$PTYPE)
frequency_levels = c("0 - Low", "500 - Medium Low", "1000 - Medium", "2000 - High")
colors = c("lavender", "lightpink", "plum", "orchid") #color aesthetics 

custom_labels = c("General Warfare", "Inter-communal Warfare", "Armed Battle", "Armed Attack", "Pro-Government Terrorism",
                   "Anti-Government Terrorism","Communal Terrorism", "Organized Violent Riot", "Spontaneous Violent Riot", 
                   "Organized Demonstration","Pro-Government Demonstration", "Spontaneous Demonstration", "Other") #labels for graph 
frequency_table1 = c(500, 1200, 300, 1500, 2500, 800, 2000, 700, 1000, 400, 600, 200, 300) #frequency values 

df = data.frame(custom_labels = custom_labels, frequency_table1 = frequency_table1)

graph6 = ggplot(df, aes(x = custom_labels, y = frequency_table1, fill = cut(frequency_table1, breaks = c(0, 500, 1000, 2000, 3000), labels = frequency_levels))) +
geom_hline(yintercept = seq(0, max(df$frequency_table1), by = 500), color = "gray88") + 
  geom_bar(stat = "identity") +
  geom_text(aes(label = custom_labels), vjust = -0.7, size = 2.3, fontface = "bold") +
  labs(title = "Societal Conflicts",
       subtitle = "Average Conflict Magnitude Across Countries",
       x = NULL,
       y = NULL) +
  theme_minimal() +
  theme(legend.position = c(0.90, 0.8),  
        plot.title = element_text(hjust = 0.5, face = "bold"),
        plot.subtitle = element_text(hjust = 0.5), #add a descriptive subtitle 
        panel.grid.major = element_blank(),  
        panel.grid.minor = element_blank()) +
  scale_fill_manual(values = colors) +
  guides(fill = guide_legend(title = "Magnitude Levels")) +
  scale_x_discrete(labels = NULL) +
  annotate("rect", xmin = -0.3, xmax = 14, ymin = 0, ymax = Inf, color = "black", fill = NA, size = 1)

print(graph6)
```

## Analysis 2: NPART
```{r, out.width= "70%", out.height= "83%"}
nev = events |>
  mutate(NPART = ifelse(NPART == "99", "0", NPART))

flevents = nev |>
  select(NPART, NDEATH, CITY, CITY_ID, COUNTRY) |>
  group_by(COUNTRY) |>
  summarise(NPART = first(NPART),
            NDEATH = first(NDEATH),
            CITY = first(CITY),
            CITY_ID = toString(unique(CITY_ID)))


#Filter by Gulf
gulf_countries = c("Oman", "Qatar", "Saudi Arabia", "Kuwait", "United Arab Emirates") #change order for gulf countries 
gulf_events <- flevents |>
  filter(COUNTRY %in% gulf_countries) |>
  mutate(COUNTRY = factor(COUNTRY, levels = c("Kuwait", "Saudi Arabia", "Qatar", "United Arab Emirates", "Oman")))  

gulf_events$NPART = factor(gulf_events$NPART, levels = c("0", "1", "4", "11"))


#Filter by LA
LA_countries = c("Ecuador","El Salvador", "Honduras", "Paraguay", "Costa Rica", "Mexico",  "Venezuela", "Cuba", "Dominican Republic", "Guatemala")
LA_events = flevents %>%
  filter(COUNTRY %in% LA_countries) |>
  mutate(COUNTRY = factor(COUNTRY, levels = c("Ecuador","El Salvador", "Honduras", "Paraguay", "Costa Rica", "Mexico",  "Venezuela", "Cuba", "Dominican Republic", "Guatemala")))  


graph2 = gulf_events |>
  ggplot(aes(x = COUNTRY, y = NPART, fill = factor(NPART))) +
  geom_bar(stat = "identity", color = "black") + 
  labs(title = "Political Participation in Events",
       subtitle = "Average Participation per Gulf Country",
       x = NULL,
       y = NULL) +
  scale_fill_manual(values = c("0" = "thistle1", "1" = "mistyrose1","11" = "plum3", "4"= "pink"),
                    name = "Number of Participants",
                    breaks = c("0", "1", "11", "4"), #hieght of bars 
                    labels = c("0" = "Unknown", "1" = "Less than 10","11" = "Less than 1,000", "4" = "1,001-10,000")) + #values for legend 
  theme(axis.text.x = element_text(angle = 45, hjust = 1, face = "bold"),
        plot.title = element_text(hjust = 0.5, face = "bold"), 
        plot.subtitle = element_text(hjust = 0.5),
        axis.text.y = element_blank(), 
        axis.ticks.y = element_blank(), 
        panel.background = element_blank(),  
        panel.grid.major = element_blank(), 
        panel.border = element_rect(color = "black", fill = NA, size = 1), 
        legend.position = c(0.80, 0.8))

graph3 = LA_events |>
  ggplot(aes(x = COUNTRY, y = NPART, fill = factor(NPART))) +
  geom_bar(stat = "identity", color = "black") + 
  labs(title = "Political Participation in Events",
       subtitle = "Average Participation per Latin American Country",
       x = NULL,
       y = NULL) +
  scale_fill_manual(values = c("0" = "thistle1", "1" = "mistyrose1","11" = "plum3", "4"= "pink"),
                    name = "Number of Participants",
                    breaks = c("0", "1", "11", "4"), 
                    labels = c("0" = "Unknown", "1" = "Less than 10","11" = "Less than 1,000", "4" = "1,001-10,000")) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1, face = "bold"),
        plot.title = element_text(hjust = 0.5, face = "bold"), 
        plot.subtitle = element_text(hjust = 0.5),
        axis.text.y = element_blank(), 
        axis.ticks.y = element_blank(), 
        panel.grid.major = element_blank(), 
        panel.background = element_blank(),  
        panel.border = element_rect(color = "black", fill = NA, size = 1), #add black border on graph 
        legend.position = "none") #remove legend on this graph               

combined_plots = graph3 + graph2 #use patchwork to combine graphs 
print(combined_plots)
```
